\documentclass[letterpaper,
            11pt]
            {moderncv}
\setcounter{tocdepth}{3}
\usepackage{geometry}
\geometry{letterpaper,left=25mm,right=20mm, bottom=30mm}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{wrapfig}
\usepackage{times} 
\usepackage{lastpage}
% Biblatex with BIBER---------------------------------
\usepackage[style=apa,sortcites=true,sorting=ydnt,backend=biber]{biblatex}
\DeclareLanguageMapping{english}{english-apa}
\addbibresource{cv-primary.bib}
\addbibresource{submitted.bib}
\addbibresource{inprep.bib}
\addbibresource{technotes.bib}
\addbibresource{talks.bib}

\moderncvtheme[blue]{classic}

\usepackage{fancyhdr}
\pagestyle{fancy}

%-----------------------------------------------------
% hebe mich als Autoren hervor
% erfordert manuelle Bearbeitung der .bib-Datei,
% ergänze bei jedem Eintrag
%     author+an = {3=highlight},
% und ersetze die "3" durch deine Position
% in der Autorenreihe
\renewcommand*{\mkbibnamegiven}[1]{%
  \ifitemannotation{highlight}
    {\textbf{#1}}
    {#1}}

\renewcommand*{\mkbibnamefamily}[1]{%
  \ifitemannotation{highlight}
    {\textbf{#1}}
    {#1}}
%-----------------------------------------------------

\rfoot{\textit{\small{\thepage/\pageref{LastPage}}}}
\name{John}{Ehrlinger, PhD}
\firstname{John}
\familyname{Ehrlinger, PhD}
\address{}{Hiram, Ohio}

\email{ehrlinj@ccf.org}

\title{Assistant Staff, Lead Data Scientist}



% Uniform name color — both parts in dark gray
\renewcommand*{\firstnamestyle}[1]{{\firstnamefont\color{color2!85!black}#1}}

% Pandoc compact-list compatibility
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Author name highlighting — \myname{} in .bib files bolds the author
\newcommand{\myname}[1]{\textbf{#1}}
\renewcommand*{\mkbibnamegiven}[1]{%
  \ifitemannotation{highlight}{\textbf{#1}}{#1}}
\renewcommand*{\mkbibnamefamily}[1]{%
  \ifitemannotation{highlight}{\textbf{#1}}{#1}}

\begin{document}

\makecvtitle


\subsection{Research Interests}\label{research-interests}

Applied statistical machine learning research conducted in close
collaboration with cardiovascular surgeons and clinicians.
Methodological focus spans random forest and ensemble methods,
clustering, deep learning, and time series analysis, with emphasis on
time-to-event and longitudinal data in cardiovascular outcomes. A
sustained focus is the translation of methodological advances into
clinical practice through open-source software development and
reproducible analytical workflows.

\subsection{Professional Experience}\label{professional-experience}

\cventry{Dec 2024--Present}{Assistant Staff, Lead Data Scientist}{Cleveland Clinic}{Cleveland, OH}{}{
  Cardiovascular Outcomes, Registries and Research (CORR) ---
  Heart, Vascular \& Thoracic Institute.
  Lead a team of data engineers and data scientists supporting cardiovascular outcomes
  research within CORR. Drive statistical methods research as applied to observational
  clinical research. Champion software engineering best practices and implement process
  improvements to optimize the departmental research pipeline.
}

\cventry{Jun 2025--Present}{Clinical Assistant Professor \textnormal{(Joint Appointment)}}{Cleveland Clinic Lerner College of Medicine}{Cleveland, OH}{}{
  Serving as faculty mentor and statistical advisor to medical students and trainees
  conducting clinical research within the Heart, Vascular \& Thoracic Institute.
  Providing guidance in statistical methods, research design, and data analysis
  for trainee-led research projects.
}

\cventry{Mar 2023--Oct 2024}{Senior Data Scientist -- Technical Lead}{Altamira Technologies}{McLean, VA}{}{
  Technical lead of data science team supporting USAF customers on secure networks.
  Built reusable Plotly-based dashboard framework; developed predictive traffic flow
  models and computer vision pipeline for vehicle volume estimation at base entry points.
}

\cventry{Jul 2015--Apr 2023}{Senior Data and Applied Scientist -- Technical Lead}{Microsoft}{Cambridge, MA}{}{
  Azure Global Commercial Industry (AGCI-AI). Technical lead for ML/AI customer
  engagements across oil \& gas (Chevron), aerospace (Rolls Royce), and medical devices
  (Stryker). Developed published Azure ML solution accelerators for predictive
  maintenance. Presented at MLADS conference four times (2016--2020).
}

\newpage

\cventry{Aug 2012--Jul 2015}{Assistant Staff / Assistant Professor of Medicine}{Cleveland Clinic}{Cleveland, OH}{}{
  Department of Quantitative Health Sciences -- Lerner Research Institute.
  Applied random survival forests to cardiovascular outcomes research.
  Initiated development of the ggRandomForests R package.
}

\cventry{2015}{Clinical Assistant Professor \textnormal{(Joint Appointment)}}{Cleveland Clinic Lerner College of Medicine}{Cleveland, OH}{}{
  Mentored medical students and fellows in statistical methods and research design
  as part of joint fellowship programs within the Cardiovascular Outcomes,
  Registries and Research (CORR) group.
}

\cventry{Aug 1999--Aug 2012}{Lead Systems Analyst: Scientific Programmer}{Cleveland Clinic}{Cleveland, OH}{}{
  Department of Thoracic and Cardiovascular Surgery -- Heart and Vascular Institute.
  Supported research computing infrastructure, statistical software development,
  and HPC operations. Completed doctoral research in Statistics at CWRU during this period.
}

\subsection{Education}\label{education}

\cventry{2011}{Doctor of Philosophy (PhD)}{Case Western Reserve University}{Cleveland, OH}{Statistics}{
  Dissertation: \textit{Regularization: Stagewise Regression and Bagging} ---
  Advisor: Hemant Ishwaran
}
\cventry{1994}{Master of Science (MS)}{Case Western Reserve University}{Cleveland, OH}{Mechanical Engineering / Aerospace Engineering}{
  Thesis: \textit{A comparison of empirical axial flow compressor models} ---
  Advisor: Frances E.\ McCaughan
}
\cventry{1993}{Bachelor of Science (BS)}{Case Western Reserve University}{Cleveland, OH}{Mechanical Engineering}{}

\subsection{Software}\label{software}

\textbf{R Packages}

\begin{itemize}
\tightlist
\item
  \emph{ggRandomForests} --- Visual exploration of random forest models.
  \url{https://github.com/ehrlinger/ggRandomForests}
\item
  \emph{boostmtree} --- Boosted multivariate trees for longitudinal
  data. \url{https://github.com/ehrlinger/boostmtree}
\item
  \emph{mixhazard} --- R port of the C code underlying the Cleveland
  Clinic Hazard SAS module. \url{https://github.com/ehrlinger/mixhazard}
\item
  \emph{hvtiPlotR} --- Publication-quality HVTI graphics.
  \url{https://github.com/ehrlinger/hvtiPlotR}
\item
  \emph{hvtiRutilities} --- Reproducible research utilities.
  \url{https://github.com/ehrlinger/hvtiRutilities}
\end{itemize}

\textbf{SAS/C Software}

\begin{itemize}
\tightlist
\item
  \emph{hazard} --- SAS and C implementation of multi-phase hazard
  analysis for time-to-event decomposition. Includes C source code, SAS
  macros and modules. (Maintainer)
  \url{https://github.com/ehrlinger/hazard}
\end{itemize}

\subsection{Honors \& Awards}\label{honors-awards}

\begin{itemize}
\tightlist
\item
  NASA Graduate Research Fellowship · 1992--1994
\item
  Cleveland Clinic Innovations Award · 2003
\end{itemize}

%----------CAVE ! --------------------------------------------
% this "\nocite{*}" command has to be set. 
% it will mark all non-cited entries for bibliography.
% Otherwise, you will get "empty bibliography"
\nocite{*}
%-------------------------------------------------------------
\printbibliography[keyword={primary}, title={Publications}, heading=subbibliography]
\printbibliography[keyword={secondary}, title={In Revision /
Submitted}, heading=subbibliography]
\printbibliography[keyword={inprep}, title={In
Preparation}, heading=subbibliography]
\printbibliography[keyword={technote}, title={Technical
Reports}, heading=subbibliography]
\printbibliography[keyword={presentation}, title={Conference
Presentations}, heading=subbibliography]

\end{document}
